---
title: "Assignment 3: Data Visualization (incarceration)"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction
Throughout the decades, social and racial issues have always been a controversial affair in the United States. Operations such as over-policing and the War on Drugs have accelerated the disproportional incarnation. Hence, the purpose of this report is to illustrate the inequality that minorities are experiencing in the justice system. To comprehend more about the subject, I have selected five variables from the dataset: the proportion of African American prison populations and the total prison population in 2010, the average jail population in the U.S. from 1970 to 2018, the proportion of White American prison populations and the total prison population in 2010, the ratio of the prison population and the general population in the United States, and the range in prison population from 1980 to 2016. 

## Summary Statistic

```{r, warning = FALSE, echo=FALSE}
library("knitr")
options(tidyverse.quiet = TRUE)
source("../source/analysis.R")

kable(summary_info)
```

To start with, the proportion of African American prison populations and the total prison population in 2010 is crucial to know about the effect of the injustice in our justice system because we can use it to compare other racial groups to understand more about the demographic in prison. The average is effective for illustrating the overall jail population across the years. In addition, we can use the proportion of White American prison populations and the total prison population in 2010 as a measure to compare other groups since they are the majority. The ratio of the prison population and the general population in the United States is practical to understand the context of the issues. Finally, the range is effective for showing the trend over time. 

## Time Trend Chart
```{r, warning = FALSE, echo=FALSE}
library("knitr")
options(tidyverse.quiet = TRUE)
source("../source/trend_data.R")

trend_data

```
Since the line chart is practical in showing the trend over time, it is suitable for this situation. The line chart shows the distribution of the prison population in the U.S by race. Starting from 2013, the African American prison population decreases yearly. The line graph leads us to reconsider the consequences of injustice in America, such as the effectiveness of over-policing, the War on Drugs, and disproportionate incarceration. 

## Variable Comparison Chart
```{r echo=FALSE, warning=FALSE}
library("knitr")
options(tidyverse.quiet = TRUE)
source("../source/two_var.R")

two_chart

```

Since a scatter plot is practical in showing the correlation between variables, it is suitable to use it in this situation. The scatterplot is calculated by extracting the data column of the African American prison population rate and White American prison population rate. Through knowing the prison population rate, we can understand the frequency of incarnation on minorities and the majority. According to the graph, starting from 2006, the prison rate for African Americans decrease yearly. It might be the result of better social programs and educational opportunities for low incomes families since, according to the [Stanford Article](https://ed.stanford.edu/in-the-media/schools-v-prisons-educations-way-cut-prison-population-op-ed-deborah-stipek) , spending more on education would shrink the prison population. Therefore, it leads us to rethink the purpose of the over-policing and War on Drugs.

## Map
```{r, warning = FALSE, echo=FALSE}
library("knitr")

options(tidyverse.quiet = TRUE)
source("../source/map.R")

map_chart

```

Map chart is practical in showing the geometrical distribution of the variables. In this analysis, I compared the total prison population in each county in Washington State to the total prison population in the United States. According to the chart, counties near the cities have a higher proportion than rural areas, such as Seattle, Spokane, Olympia, and Tacoma. It may be because the compact land use in the city makes it easier to over-policing. This chart clearly shows us the places where over-policing may happen. 

## References
- Stipek , D., &amp; Hanson, K. (2014, May 19). Schools v. prisons: Education's the way to cut prison population (op-ed by Deborah Stipek). Stanford Graduate School of Education. Retrieved March 11, 2022, from https://ed.stanford.edu/in-the-media/schools-v-prisons-educations-way-cut-prison-population-op-ed-deborah-stipek 
